---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    fig_caption: yes
    highlight: tango
    keep_md: yes
    theme: cerulean
date: "20 December 2015"
---
**Goal**: To create a *dynamic document*, in which thourougly explains the steps
taken for analysis on the activity data. Futhermore, the objective includes 
"...[a written] report that answers the questions detailed below. Ultimately, 
[we] will need to complete the entire assignment in a single R markdown document
that can be processed by knitr and be transformed into an HTML file." 
(*Coursera, Roger D. Peng*)

1. Loading and preprocessing the data

```{r}
# Step 1: Set WD to cloned repository (located on the Computer)
setwd("~/Desktop/RepData_PeerAssessment1")

# Step 2: Unzip file and Read the data into descriptive object
activity <- unzip("activity.zip", "activity.csv")
active_data <- read.csv(activity)

# Step 3: Look and Clean the Data, remove "NA's"
str(active_data) # View data Structure
active_data<- na.omit(active_data) # Removing "NA's"

# Step 4: Re-Peak at the Data Frame to ensure no "NA's"
str(active_data)
head(active_data)

# Step 5: Use dplyr package for easier manipulation of data frame
library(dplyr) 
active_data <- tbl_df(active_data) # Create dplyr data frame

# Now, we can see we have a "clean" data frame with 3 variables and 15264 
# observations! We can now move onto the second part of the analysis!
```

2. What is mean total number of steps taken per day?
```{r}
# Step 1: Load ggplot2 for graphing
library(ggplot2) 

# Step 2: Group Steps by Day, specifically the variable "Date"
str(active_data$date) # See what type of Variable "Date" is in data frame

# Step 3: Histogram of Steps per Day
steps<- ggplot(active_data, aes(x=factor(date), y=steps, fill=date, width=.75))
steps <- steps+ geom_bar(color="black", stat="identity")+ scale_fill_hue(name="date")
steps <- steps+ xlab("Date") + ylab("Amount Steps") + theme(axis.text.x= element_text(angle=90, size=6)) 
steps <- steps + guides(fill=guide_legend(ncol=2))
print(steps)

# Step 4: Histogram(interval) of Steps each Day
day_step<- aggregate(steps ~ date, active_data, sum) # Group by Date and then Sum the Steps
step <- hist(day_step$steps, xlab= "Number of Steps", main="Total Amount of Steps Each Day!", col="purple")
print(step) 
# This Graph shows number of Steps each day as a larger interval histogram 
# instead of individual dates interval.

# Step 5: Calculate table with mean steps by date
group_date <- group_by(active_data, date) # Group data frame by date
avg_steps_bydate <- summarise(group_date, Mean_Steps= mean(steps, na.rm=TRUE)) 
# Calculated mean by grouped "date" above

```

```{r out.width=c('900px', '900px'), fig.show='hold'}
# Step 6: Print Out visual descriptive, put table and Historgram to represent
# the Mean Number of Steps per day(date)
tbl_df(avg_steps_bydate) # View the Data frame with Date and Average/Mean Steps

# Step 7: Add layers to graph
graph<- ggplot(avg_steps_bydate, aes(x=factor(date), y=Mean_Steps, fill=date, width=1))
graph <- graph+ geom_bar(color="black", stat="identity")+ scale_fill_hue(name="date")
graph <- graph+ xlab("Date") + ylab("Mean Steps") + theme(axis.text.x= element_text(angle=90, size=6)) 
graph <- graph+ ggtitle("Average Steps per Day") + theme(legend.position="botton")
graph <- graph+ theme(plot.title = element_text(size=10, face="bold", vjust=2))
graph <- graph + theme(legend.background = element_rect()) 
graph <- graph + theme(legend.background = element_rect(fill="gray90", size=1, linetype="dotted"))
graph <- graph + guides(fill=guide_legend(ncol=10))
print(graph)

# Step 8: Histogram(interval) of Average Steps per Day
avg_step<- aggregate(steps ~ date, active_data, mean) # Group by Date and then Sum the Steps
avg_step <- hist(avg_step$steps, xlab= "Mean Number of Steps", main="Average Amount of Steps per Each Day!", col="orange")
print(avg_step) # This Graph shows number of Steps each day as a interval histogram 
# instead of individual dates

# Step 7: Now for the median we do the same as above
group_date <- group_by(active_data, date) # Group data frame by date
median_steps <- summarise(group_date, Median_Steps= median(steps))
```

3. What is the average daily activity pattern?



4. Imputing missing values



5. Are there differences in activity patterns between weekdays and weekends?
